KidsAIR: initial exposure analysis

### 2020-03-26 demographics_summary.Rmd
data cleaning/merging updates

---
title: "KidsAIR: household demographics summary"
author: "Ethan Walker"
date: "Started 27 Jan 2020, Updated 13 March 2020"
output: 
  powerpoint_presentation
---

@@

# Summarize demographics data
```{r, include=FALSE}
# Summarize numeric vars
kids_dems <- demographics_filtered %>%  
  mutate(total_residents = as.numeric(total_residents),
         residents_under_five = as.numeric(residents_under_five)) %>% 
  distinct(home, .keep_all = TRUE) %>%  
  group_by(area) %>% 
  summarize("Mean" = mean(total_residents, na.rm = TRUE), 
            "SD" = sd(total_residents, na.rm = TRUE),
            "min" = min(total_residents, na.rm = TRUE), 
            "max" = max(total_residents, na.rm = TRUE))
kable(kids_dems, digits = 1, align = "c")

@@

demographics_filtered <- demographics_clean %>% 
  filter(winter_id == 1) %>% 
  filter(!is.na(hispanic_child)) %>% 
  mutate(income = factor(income,
                         levels = c("Less than $20,000", "$20,000 to $29,999",
                                    "$30,000 to $39,999", "$40,000 to $49,999",
                                    "$50,000 to $74,999", "$75,000 to $99,999",
                                    "$100,000 or more")),
         education = factor(education,
                            levels = c("Less than high school",
                                       "High school diploma or GED",
                                       "Some college",
                                       "College degree")),
         total_residents = as.numeric(total_residents),
         race_parent = factor(race_parent,
                              levels = c("American Indian/Alaskan Native",
                                         "Asian", "White",
                                         "More than one race"),
                              labels = c("AI/AN", "Asian", "White", "Multiple")),
         race_child = factor(race_child,
                              levels = c("American Indian/Alaskan Native",
                                         "White", "More than one race"),
                              labels = c("AI/AN", "White", "Multiple")))
```

@@

  demographics_plot <- data %>% 
  filter(winter_id == 1) %>% 
  select(area, var) %>% 
  rename(plot_var = var) %>% 
  filter(!is.na(plot_var)) %>% 
  ggplot() +
    geom_bar(aes(plot_var, fill = area)) +
    theme_minimal() +
    theme(axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 40, vjust = 0.7, hjust = 0.7),
          axis.text.y = element_text(size = 16, colour = "black"),
          title = element_text(size = 18, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          panel.grid = element_blank()) +
    labs(title = title_label) +
    ylab(y_label) +
    scale_fill_manual(values = jv_palette)
  demographics_plot

@@

```{r}
demographics_function(demographics_filtered, "residents_under_five", "Residents under 5", "count")
```

>>>>

---
title: 'KidsAIR: household demographics summary'
author: "Ethan Walker"
date: "Started 27 Jan 2020, Updated 25 March 2020"
output:
  beamer_presentation: default
  powerpoint_presentation: default
---

@@

```{r, include=FALSE}
# Summarize numeric vars
kids_dems <- demographics_filtered %>%  
  mutate(total_residents = as.numeric(total_residents),
         residents_under_five = as.numeric(residents_under_five)) %>% 
  distinct(home, .keep_all = TRUE) %>%  
  group_by(area) %>% 
  summarize("Mean" = mean(total_residents, na.rm = TRUE), 
            "SD" = sd(total_residents, na.rm = TRUE),
            "min" = min(total_residents, na.rm = TRUE), 
            "max" = max(total_residents, na.rm = TRUE))
kable(kids_dems, digits = 1, align = "c")

@@

demographics_filtered <- demographics_clean %>% 
  filter(winter_id == 1) %>% 
  filter(residents_under_five != 0) %>% 
  mutate(income = factor(income,
                         levels = c("Less than $20,000", "$20,000 to $29,999",
                                    "$30,000 to $39,999", "$40,000 to $49,999",
                                    "$50,000 to $74,999", "$75,000 to $99,999",
                                    "$100,000 or more")),
         education = factor(education,
                            levels = c("Less than high school",
                                       "High school diploma or GED",
                                       "Some college",
                                       "College degree")),
         total_residents = as.numeric(total_residents),
         total_residents = if_else(total_residents == 1, 2, total_residents),
         race_parent = factor(race_parent,
                              levels = c("American Indian/Alaskan Native",
                                         "Asian", "White",
                                         "More than one race"),
                              labels = c("AI/AN", "Asian", "White", "Multiple")),
         race_child = factor(race_child,
                              levels = c("American Indian/Alaskan Native",
                                         "White", "More than one race"),
                              labels = c("AI/AN", "White", "Multiple")),
         age_child_nov = factor(age_child_nov,
                                levels = c("Less than 1", "1", "2", "3", "4")))
```

@@

  demographics_plot <- data %>% 
  filter(winter_id == 1) %>% 
  select(area, var) %>% 
  rename(plot_var = var) %>% 
  filter(!is.na(plot_var)) %>% 
  ggplot() +
    geom_bar(aes(plot_var, fill = area)) +
    theme_minimal() +
    theme(axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 40, vjust = 0.7, hjust = 0.7),
          axis.text.y = element_text(size = 16, colour = "black"),
          title = element_text(size = 20, colour = "black"),
          legend.text = element_text(size = 18, colour = "black"),
          legend.title = element_text(size = 18, colour = "black"),
          axis.title.y = element_text(size = 18, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          panel.grid = element_blank()) +
    labs(title = title_label) +
    ylab(y_label) +
    scale_fill_manual(values = jv_palette)
  demographics_plot

@@

```{r}
demographics_function(demographics_filtered, "residents_under_five", 
                      "Residents under 5 years of age", "count")
```

### 2020-03-26 exposure_initial_analysis.Rmd
data cleaning/merging updates

---
title: "KidsAIR: initial exposure analysis"
author: "Ethan Walker"
date: "Started 2 Feb 2020, Updated 20 March 2020"
output: pdf_document
---

@@

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(lme4)
library(lmerTest)
```

# Load individual datasets
```{r}
pm_clean <- read_rds("Output/pm_clean.rds")
sums_clean <- read_rds("Output/sums_clean.rds")

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
demographics_clean <- read_rds("Output/demographics_clean.rds")
moisture_clean <- read_rds("Output/moisture_clean.rds")
indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
stove_use_clean <- read_rds("Output/stove_use_clean.rds")
stove_grades_clean <- read_rds("Output/stove_grades_clean.rds")
home_char_clean <- read_rds("Output/home_char_clean.rds")
home_act_clean <- read_rds("Output/home_act_clean.rds")

# Mean PM for each placebo home over winter 1
pm_home_mean <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  #filter(pm < 500) %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  group_by(area, home) %>% 
  summarize(mean_pm = mean(pm, na.rm = TRUE))

exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long.rds")

exposure_analysis_data_short <- read_rds("Output/exposure_analysis_data_short.rds") 
```

# Demographics
```{r}
filter_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  distinct(home, .keep_all = TRUE) %>% 
  filter(!is.na(pm_mean))

@@

table(filter_data$residents_smoke)
median(filter_data$home_furry, na.rm = TRUE)

pm_demographics_results <- lm(log(pm_mean) ~ residents_smoke, data = filter_data)
summary(pm_demographics_results)
plot(pm_demographics_results)
```

# Home characteristics
```{r}
filter_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  distinct(home, .keep_all = TRUE) %>% 
  filter(!is.na(pm_mean))

@@

table(filter_data$home_insert)
median(filter_data$home_furry, na.rm = TRUE)

pm_home_char_results <- lm(log(pm_mean) ~ home_insert, data = filter_data)
summary(pm_home_char_results)
plot(pm_home_char_results)
```

# Stove use
```{r}
filter_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  distinct(home, .keep_all = TRUE) %>% 
  filter(!is.na(pm_mean))

@@

table(filter_data$home_insert)
median(filter_data$wood_cords, na.rm = TRUE)

pm_stove_use_results <- lm(log(pm_mean) ~ wood_collect_method_2level, data = filter_data)
summary(pm_stove_use_results)
plot(pm_stove_use_results)
pm_stove_use_plot <- pm_stove_use %>% 
  ggplot() + 
    geom_boxplot(aes(burn_level, mean_pm))
pm_stove_use_plot
```

# Stove grades
```{r}
filter_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  distinct(home, .keep_all = TRUE) %>% 
  filter(!is.na(pm_mean))

@@

table(filter_data$stove_grade_3level)

pm_stove_grades_results <- lm(log(pm_mean) ~ stove_grade_3level, data = filter_data)
summary(pm_stove_grades_results)
anova(pm_stove_grades_results)
plot(pm_stove_grades_results)
pm_stove_grades_plot <- pm_stove_grades %>% 
  ggplot() + 
    geom_boxplot(aes(stove_grade, mean_pm))
pm_stove_grades_plot
```

# Home activity
```{r}
filter_data <- exposure_analysis_data_long %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(home) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  filter(!is.na(pm_mean_sampling_day))

@@

table(filter_data$home_act_sweep)

pm_home_act_results <- lmer(log(pm_mean_sampling_day) ~ home_act_sweep + 
                              (1 | home), data = filter_data)
summary(pm_home_act_results)
plot(pm_home_act_results, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_home_act_results), main = "QQ Plot")
```

# Wood moisture content
```{r}
filter_data <- exposure_analysis_data_long %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  filter(!is.na(moisture_ave)) %>% 
  select(sample_date, pm_first_datetime, area, home, 
         moisture_ave, pm_mean_sampling_period) %>% 
  separate(pm_first_datetime, c("pm_first_date", "first_time"), sep = " ") %>% 
  group_by(area, home) %>% 
  mutate(moisture_date = ymd(sample_date),
         first_date = ymd(pm_first_date),
         time_diff = moisture_date - first_date,
         time_diff = as.numeric(abs(time_diff))) %>% 
  group_by(home) %>% 
  mutate(time_diff_rank = if_else(time_diff == min(time_diff, na.rm = TRUE), 1, 0)) %>%
  ungroup() %>% 
  filter(time_diff_rank == 1) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  filter(!is.na(pm_mean_sampling_period))

# Variables to use:
## moisture_ave, mean_moisture_5

table(filter_data$moisture_cat)
median(filter_data$moisture_ave, na.rm = TRUE)

pm_moisture_results <- lm(log(pm_mean_sampling_period) ~ 
                            moisture_ave, data = filter_data)
summary(pm_moisture_results)
plot(pm_moisture_results)
```

# Indoor temp/Rh
```{r}
pm_temp_rh <- pm_home_mean %>% 
  left_join(indoor_temp_rh_clean, by = c("area", "home")) %>% 
  filter(winter_id == 1) %>% 
  mutate(temp_max_5 = temp_indoor_max/5,
         temp_min_5 = temp_indoor_min/5,
         rh_max_5 = rh_indoor_max/5,
         rh_min_5 = rh_indoor_min/5,
         temp_max_cat = cut(temp_indoor_max, breaks = c(0, 30, 45),
                                labels = c("<30", "30+"),
                                 right = FALSE),
         temp_min_cat = cut(temp_indoor_min, breaks = c(0, 16.5, 50),
                                labels = c("<16.5", "16.5+"),
                                 right = FALSE),
         rh_max_cat = cut(rh_indoor_max, breaks = c(0, 40, 100),
                                labels = c("<40", "40+"),
                                 right = FALSE),
         rh_min_cat = cut(rh_indoor_min, breaks = c(0, 20, 100),
                                labels = c("<20", "20+"),
                                 right = FALSE))

table(pm_temp_rh$rh_min_cat)
median(pm_temp_rh$rh_indoor_min, na.rm = TRUE)
range(pm_temp_rh$rh_indoor_max, na.rm = TRUE)

pm_home_act_results <- lm(log(mean_pm) ~ rh_min_cat, data = filter_data)
summary(pm_home_act_results)
plot(pm_home_act_results)
```

# SUMs - experimenting with several methods of looking at associations
```{r}
#### Join PM dataset full SUMs dataset by SUMs sampling times

pm_join <- pm_clean %>% 
  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
  group_by(area, home) %>% 
  arrange(sample_datetime) %>% 
  mutate(pm_last_datetime = last(sample_datetime),
         pm_first_datetime = first_datetime_new,
         sample_datetime = round_date(sample_datetime, unit = "minute"),
         pm_rolling_20 = rollmean(pm, 20, fill = "extend")) %>% 
  ungroup() %>% 
  arrange(area, home, sample_datetime)

sums_join <- sums_clean %>% 
  mutate(sample_datetime = ymd_hms(datetime_sums)) %>% 
  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
  mutate(sample_datetime = round_date(sample_datetime, unit = "minute")) %>% 
  ungroup() %>% 
  arrange(area, home, sample_datetime)

pm_sums_joined <- sums_join %>% 
  left_join(pm_join, by = c("area", "home", "sample_datetime")) %>% 
  filter(pm_rolling_20 < 1000) %>% 
  filter(!is.na(temp_c_sums)) %>% 
  filter(!is.na(pm_rolling_20)) %>% 
  mutate(temp_c_sums_5 = temp_c_sums/5)

pm_sums_count <- pm_sums_joined %>% 
  count(home)

cor.test(pm_sums_joined$temp_c_sums, pm_sums_joined$pm_rolling_20, method = "spearman")
pm_sums_model <- lmer(log(pm_rolling_20) ~ temp_c_sums_5 + (1 | home), data = pm_sums_joined)

summary(pm_sums_model)
exp((0.146291)-1)*100
plot(pm_sums_model, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_model), main = "QQ Plot")

pm_sums_plot <- pm_sums_joined %>% 
  ggplot() + 
    geom_point(aes(log(pm_rolling_20), temp_c_sums)) +
    geom_smooth(aes(log(pm_rolling_20), temp_c_sums)) +
    theme_minimal()
pm_sums_plot

#### Join PM dataset with SUMs dataset by sampling day

pm_join <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  mutate(pm = if_else(pm > 500, 500, pm),
         pm_mean = mean(pm, na.rm = TRUE)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(pm_mean_daily = mean(pm, na.rm = TRUE)) %>% 
  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
  group_by(area, home, sampling_day) %>% 
  arrange(sample_datetime) %>% 
  mutate(pm_last_datetime = last(sample_datetime),
         pm_first_datetime = first(sample_datetime)) %>% 
  ungroup() %>% 
  arrange(area, home, sample_datetime) %>% 
  select(area, home, home_winter_id, dtid, day_of_week, pm_total_observations,
         pm_mean, sampling_day, pm_mean_daily, pm_first_datetime, pm_last_datetime) %>% 
  group_by(area, home) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home, sampling_day)

sums_join <- sums_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  select(area, home, winter_id, ibid, treatment, datetime_sums, temp_c_sums,)

pm_sums_joined <- sums_join %>% 
  left_join(pm_join, by = c("area", "home")) %>% 
  arrange(area, home, datetime_sums) %>% 
  mutate(filter_var = if_else(datetime_sums >= pm_first_datetime &
                              datetime_sums <= pm_last_datetime, 1, 0)) %>% 
  filter(filter_var == 1) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(temp_perc_25 = if_else(temp_c_sums >= 30, 1, 0)) %>% 
  arrange(area, home, sampling_day) %>% 
  mutate(sums_mean_daily = mean(temp_c_sums, na.rm = TRUE),
         sums_mean_daily_5 = sums_mean_daily/5) %>% 
  group_by(area, home, home_winter_id, sampling_day, pm_mean, 
           pm_mean_daily, sums_mean_daily) %>% 
  summarize(temp_perc_25 = (sum(temp_perc_25)/n()*100)) %>% 
  mutate(temp_perc_25_cat = cut(temp_perc_25, breaks = c(0, 10, 200),
                                labels = c("< 50%", "50% +"),
                                 right = FALSE)) %>%
  distinct(pm_mean_daily, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home, sampling_day)

pm_sums_count <- pm_sums_joined %>% 
  count(home)
median(pm_sums_joined$temp_perc_25)
table(pm_sums_joined$temp_perc_25_cat)

cor.test(pm_sums_joined$sums_mean_daily, pm_sums_joined$pm_mean_daily, method = "spearman")
pm_sums_model <- lmer(pm_mean_daily ~ temp_perc_25_cat + (1 | home), data = pm_sums_joined)
pm_sums_model <- lmer(log(pm_mean_daily) ~ temp_perc_25_cat + (1 | home), data = pm_sums_joined)

summary(pm_sums_model)
exp((0.146291)-1)*100
plot(pm_sums_model, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_model), main = "QQ Plot")

#### Join PM dataset with SUMs events datasets

# Mean PM for each placebo home over winter 1
pm_home_mean_date <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  separate(pm_datetime_new, c("date_sums", "pm_time"), sep = " ", remove = TRUE) %>% 
  group_by(area, home, date_sums) %>% 
  summarize(mean_pm = mean(pm, na.rm = TRUE)) %>% 
  group_by(area, home, date_sums) %>% 
  distinct(date_sums, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home, date_sums)

sums_temp_change <- sums_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home, date_sums) %>% 
  mutate(time_diff = (lead(datetime_sums) - datetime_sums)/60) %>% 
  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
         lead_temp = if_else(time_diff > 25, 999, lead_temp)) %>% 
  replace_with_na(replace = list(lead_temp = 999)) %>% 
  mutate(temp_diff = lead_temp - temp_c_sums,
         temp_diff_abs = abs(lead_temp - temp_c_sums),
         # change temp in the following line
         temp_diff_check = if_else(temp_diff >= 8, 1, 0),
         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
                              temp_diff_check),
         heat_event = if_else(is.na(temp_diff_check), 1,
                              heat_event)) %>% 
  ungroup() %>% 
  mutate(date_sums = as.character(date_sums)) %>% 
  group_by(area, home, date_sums) %>% 
  summarize(sums_events = sum(heat_event, na.rm = TRUE)) %>% 
  ungroup() %>% 
  arrange(area, home, date_sums)

pm_sums_events_joined <- pm_home_mean_date %>% 
  left_join(sums_temp_change, by = c("area", "home", "date_sums")) %>% 
  arrange(area, home, date_sums) %>% 
  #mutate(sums_events = factor(sums_events))
  mutate(sums_events_cat = cut(sums_events, breaks = c(0, 2, 20),
                           labels = c("<2", "2+"),
                                 right = FALSE))

median(pm_sums_events_joined$sums_events, na.rm = TRUE)
table(pm_sums_events_joined$sums_events_cat)

pm_sums_model <- lmer(mean_pm ~ sums_events_cat + (1 | home), data = filter_data)
pm_sums_model <- lmer(log(mean_pm) ~ sums_events_cat + (1 | home), data = filter_data)

@@

pm_sums_plot <- pm_sums_events_joined %>% 
  filter(!is.na(sums_events)) %>% 
  ggplot() + 
    geom_boxplot(aes(sums_events_cat, log(mean_pm))) +
    theme_minimal()
pm_sums_plot

```

>>>>

---
title: "KidsAIR: initial exposure analysis"
author: "Ethan Walker"
date: "Started 2 Feb 2020, Updated 26 March 2020"
output: pdf_document
---

@@

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(broom)
library(broom.mixed)
library(zoo)
library(lme4)
library(lmerTest)
```

# Load individual datasets
```{r}
exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long_new.rds")

exposure_analysis_data_medium <- read_rds("Output/exposure_analysis_data_medium.rds")

exposure_analysis_data_short <- read_rds("Output/exposure_analysis_data_short.rds")

sums_rolling_pm_data <- read_rds("Output/sums_rolling_pm_data.rds")
```

# Demographics
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

@@

table(analysis_data$residents_smoke)
median(analysis_data$home_furry, na.rm = TRUE)

pm_demographics_results <- lm(log(pm_mean_sampling_period) ~ residents_smoke, 
                              data = analysis_data)
summary(pm_demographics_results)
tidy(pm_demographics_results)
plot(pm_demographics_results)
```

# Home characteristics
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

@@

table(analysis_data$home_insert)
median(analysis_data$home_furry, na.rm = TRUE)

pm_home_char_results <- lm(log(pm_mean_sampling_period) ~ home_insert, 
                           data = analysis_data)
summary(pm_home_char_results)
tidy(pm_home_char_results)
plot(pm_home_char_results)
```

# Stove use
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

@@

table(analysis_data$burn_level_3level)
median(analysis_data$wood_cords, na.rm = TRUE)

pm_stove_use_results <- lm(log(pm_mean_sampling_period) ~ wood_collect_method_2level, 
                           data = analysis_data)
summary(pm_stove_use_results)
tidy(pm_stove_use_results)
plot(pm_stove_use_results)
pm_stove_use_plot <- pm_stove_use %>% 
  ggplot() + 
    geom_boxplot(aes(burn_level, mean_pm))
pm_stove_use_plot
```

# Stove grades
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

@@

table(analysis_data$stove_grade_3level)

pm_stove_grades_results <- lm(log(pm_mean_sampling_period) ~ stove_grade_3level, 
                              data = analysis_data)
summary(pm_stove_grades_results)
tidy(pm_stove_grades_results)
anova(pm_stove_grades_results)
plot(pm_stove_grades_results)
pm_stove_grades_plot <- pm_stove_grades %>% 
  ggplot() + 
    geom_boxplot(aes(stove_grade, mean_pm))
pm_stove_grades_plot
```

# Home activity
```{r}
analysis_data <- exposure_analysis_data_medium %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  ungroup()

@@

table(analysis_data$home_act_sweep)

pm_home_act_results <- lmer(log(pm_mean_sampling_day) ~ home_act_sweep + 
                              (1 | home), data = analysis_data)
summary(pm_home_act_results)
tidy(pm_home_act_results)
plot(pm_home_act_results, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_home_act_results), main = "QQ Plot")
```

# Wood moisture content
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE),
         moisture_winter_2level = cut(mean_moisture_winter, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE))

# Variables to use:
## moisture_closest, moisture_closest_2level
## mean_moisture_winter, moisture_winter_2level

table(analysis_data$moisture_closest_2level)
median(analysis_data$mean_moisture_winter, na.rm = TRUE)

pm_moisture_results <- lm(log(pm_mean_sampling_period) ~ 
                            moisture_winter_2level, data = analysis_data)
summary(pm_moisture_results)
plot(pm_moisture_results)
```

# Indoor temp/RH
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(temp_max_2level = cut(temp_indoor_max, breaks = c(0, 30, 100),
                                 labels = c("<30", "30+"),
                                 right = FALSE),
         temp_min_2level = cut(temp_indoor_min, breaks = c(0, 16.5, 100),
                                 labels = c("<16.5", "16.5+"),
                                 right = FALSE),
         rh_max_2level = cut(rh_indoor_max, breaks = c(0, 40, 100),
                                 labels = c("<40", "40+"),
                                 right = FALSE),
         rh_min_2level = cut(rh_indoor_min, breaks = c(0, 20, 100),
                                 labels = c("<20", "20+"),
                                 right = FALSE))

# Variables to use:
## rh_min_5, rh_max_5, temp_min_5, temp_max_5
## temp_max_2level, temp_min_2level,rh_max_2level ,rh_min_2level

table(analysis_data$temp_max_2level)
median(analysis_data$rh_indoor_min, na.rm = TRUE)

pm_temp_rh_results <- lm(log(pm_mean_sampling_period) ~ temp_max_2level, 
                         data = analysis_data)
summary(pm_temp_rh_results)
tidy(pm_temp_rh_results)
plot(pm_temp_rh_results)
```

# SUMs - by sampling day
```{r}
analysis_data <- exposure_analysis_data_medium %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  mutate(sums_mean_sampling_day_5 = sums_mean_sampling_day/5,
         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
                                 labels = c("<50", "50+"),
                                 right = FALSE),
         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
                                 labels = c("<35", "35+"),
                                 right = FALSE),
         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
                                 labels = c("<10", "10+"),
                                 right = FALSE)) %>% 
  ungroup()

# Variables to use:
## sums_mean_sampling_day_5, temp_perc_25_2level
## temp_perc_27_2level, temp_perc_30_2level

table(analysis_data$temp_perc_25_2level)
median(analysis_data$temp_perc_30, na.rm = TRUE)

pm_sums_results <- lmer(log(pm_mean_sampling_day) ~ temp_perc_25_2level + 
                              (1 | home), data = analysis_data)
summary(pm_sums_results)
tidy(pm_sums_results)
plot(pm_sums_results, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_results), main = "QQ Plot")
```

# SUMs - by sampling period
```{r}
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  mutate(sums_mean_sampling_period_5 = sums_mean_sampling_period/5,
         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
                                 labels = c("<50", "50+"),
                                 right = FALSE),
         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
                                 labels = c("<35", "35+"),
                                 right = FALSE),
         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
                                 labels = c("<10", "10+"),
                                 right = FALSE)) %>% 
  ungroup()

# Variables to use:
## sums_mean_sampling_period_5, temp_perc_25_2level
## temp_perc_27_2level, temp_perc_30_2level

table(analysis_data$temp_perc_25_2level)
median(analysis_data$temp_perc_30, na.rm = TRUE)

pm_sums_results <- lm(log(pm_mean_sampling_period) ~ sums_mean_sampling_period_5, 
                      data = analysis_data)
summary(pm_sums_results)
tidy(pm_sums_results)
plot(pm_sums_results, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_results), main = "QQ Plot")
```

# SUMs - mean PM per SUMs interval
```{r}
analysis_data <- sums_rolling_pm_data %>% 
  mutate(temp_c_sums_5 = temp_c_sums/5)

cor.test(analysis_data$temp_c_sums, analysis_data$pm_rolling_20, method = "spearman")
pm_sums_model <- lmer(log(pm_rolling_20) ~ temp_c_sums_5 + (1 | home), 
                      data = analysis_data)

summary(pm_sums_model)
tidy(pm_sums_model)
exp((0.146291)-1)*100
plot(pm_sums_model, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_model), main = "QQ Plot")

pm_sums_plot <- analysis_data %>% 
  ggplot() + 
    geom_point(aes(log(pm_rolling_20), temp_c_sums)) +
    geom_smooth(aes(log(pm_rolling_20), temp_c_sums)) +
    theme_minimal()
pm_sums_plot
```

# SUMs - events per sampling day - KEEP WORKING ON THIS
```{r}
sums_temp_change <- sums_rolling_pm_data %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home, sample_date) %>% 
  distinct(datetime_sums, .keep_all = TRUE) %>% 
  mutate(time_diff = (lead(datetime_sums) - datetime_sums)/60) %>% 
  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
         lead_temp = if_else(time_diff > 25, 999, lead_temp)) %>% 
  replace_with_na(replace = list(lead_temp = 999)) %>% 
  mutate(temp_diff = lead_temp - temp_c_sums,
         temp_diff_abs = abs(lead_temp - temp_c_sums),
         # change temp in the following line
         temp_diff_check = if_else(temp_diff >= 8, 1, 0),
         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
                              temp_diff_check),
         heat_event = if_else(is.na(temp_diff_check), 1,
                              heat_event)) %>% 
  ungroup() %>% 
  mutate(sample_date = as.character(sample_date)) %>% 
  group_by(area, home, sample_date) %>% 
  mutate(sums_events = sum(heat_event, na.rm = TRUE),
         sums_events_cat = cut(sums_events, breaks = c(0, 5, 50),
                           labels = c("<4", "4+"),
                                 right = FALSE)) %>% 
  ungroup() %>% 
  arrange(area, home, sample_date)

median(sums_temp_change$sums_events, na.rm = TRUE)
table(sums_temp_change$sums_events_cat)

pm_sums_model <- lmer(mean_pm ~ sums_events + (1 | home), data = sums_temp_change)
pm_sums_model <- lmer(log(pm_mean_daily) ~ sums_events + (1 | home), data = sums_temp_change)

@@

pm_sums_plot <- pm_sums_events_joined %>% 
  filter(!is.na(sums_events)) %>% 
  ggplot() + 
    geom_boxplot(aes(sums_events_cat, log(mean_pm))) +
    theme_minimal()
pm_sums_plot
```
